### 2020-12-22 manuscript/abstract.Rmd
Submitted to AEE

Ground-dwelling arthropods can be important generalist predators in agroecosystems, and can use non-crop features as overwintering habitats. However, it is unclear which types of landscape features constitute useful non-crop habitat, and at what spatial scale organisms gather resources. Additionally, the same landscape feature may act as a source or a destination for arthropods at different times of the year, but this is rarely considered. We modeled the abundance of four common species of Canadian prairie arthropods caught in a set of 198 in-field and roadside pitfall traps (June to August of 2017). Functional regression was used in order to simultaneously consider both the habitat preferences and the timing of movement from the land cover classes. _Pterostichus melanarius_ (Coleoptera: Carabidae) and _Pardosa moesta_ (Araneae: Lycosidae) were attracted to canola (_Brassica napus_) during the early summer, then dispersed to grasslands, wetlands, and grassy road margins at the end of the summer. In particular, _Pterostichus melanarius_ aggregated in canola early in the growing season, suggesting that its role in suppressing crop pests may be underestimated. _Pardosa distincta_ (Araneae: Lycosidae) and _Phalangium opilio_ (Opiliones: Opilionidae) showed weak patterns of seasonal migration, and were more influenced by large-scale geographic patterns rather than landscape composition. Our results suggest that predatory arthropods migrate into canola crops during the early summer, and that grasslands and wetlands act as seasonal reservoir habitats. Farmers and land managers should consider preserving existing habitat in order to maintain pest-control services across the season.