### 2021-05-25 manuscript/supplemental2.Rmd

\newpage

# Appendix B: R Code {-}

```{r modelFormula, eval=FALSE}

#Code to run functional regression using gam function from mgcv 1.8-35

gam(count ~ offset(log(trapdays)) + trapLoc + s(day) + s(E,N)+ 
      #Canola
      s(distanceMatrix,by=Canola) + #Functional regression of distance
      s(dayMatrix,by=Canola) + #FR of time
      ti(distanceMatrix,dayMatrix,by=Canola) + #FR of distance and time
      #Pasture
      s(distanceMatrix,by=Pasture)+ #Functional regression of distance
      s(dayMatrix,by=Pasture) + #FR of time 
      ti(distanceMatrix,dayMatrix,by=Pasture) + #FR of distance and time
      #Other land cover classes here
      #...
      #...
      #...
      s(distanceMatrix,by=Roads) + 
      s(dayMatrix,by=Roads)+ 
      ti(distanceMatrix,dayMatrix,by=Roads),
    family= 'nb'
)
# where:
#
# count = N x 1 vector of activity density from traps (count)
# trapdays = N x 1 vector of number of days that the trap was installed for (numeric)
# trapLoc = N x 1 vector of trap location type (factor)
# day = N x 1 vector of day of year that the trap was installed on (numeric)
# E,N = N x 1 vector(s) of Easting and Northing (numeric)
# distanceMatrix = N x M matrix of distances, where the Nth row contains a vector
#     of distances to the annuli (numeric)
# dayMatrix = N x M matrix of days, where the Nth row contains a vector of days 
#     that spans the range of days of the year (numeric)
# Canola, Pasture, ... = N x M matrix of cover measurements, where the Nth row
#     contains a vector of proportion cover measurements within each annulus (numeric)
# 'nb' = Negative binomial family
```